Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2021-10-18 doc/AMS-15/AMS-15_poster.Rmd
Improve figure captions and formatting. Add and rearrange text. Remove extra figures.

# Results

@@

### Flow vs current

@@

```{r cur-flow-os, fig.height=4, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  labs(title = "Vial dilution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic()
```

@@

```{r blanks, fig.height=5, dev='svg'}
cons_df %>% 
  filter(fm_consensus < 0.1) %>% 
  ggplot(aes(he12C, norm_ratio)) +
  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio  + sig_norm_ratio)) +
  geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C  + sig_he12C)) +
  geom_point(aes(color = wheel), size = 3) +
  #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
  labs(title = "Blanks",
       x = "12C Current (μA)",
       y = "Fraction modern") +
  theme_classic() +
  theme(legend.position = c(0.82, 0.76),
        legend.background = element_rect(fill = "white", color = "black")) 
```

## Results: Method precision and accuracy

@@

cons_df %>% 
ggplot(aes(fm_consensus, fm_diff, color = Name)) + 
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = mean_diff, color = "blue") +
  geom_hline(yintercept = mean_diff + sd_diff, color = "lightblue") +
  geom_hline(yintercept = mean_diff - sd_diff, color = "lightblue") +
  geom_smooth(method = "lm",  se = FALSE) + 
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr), 
                  position = position_dodge2(width = 0.1), 
                  size = 0.2) + 
  labs(subtitle = "Blank corrected", x = "Fm expected (points jittered for clarity)", 
       y = "Fm difference (cons - measured)")
```

@@

To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method compared to the same samples measured using the traditional graphite method. We found the measurements to be in good agreement, with no significant bias and errors averaging XXX compared with XXX for graphite measurements.

@@

combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
  mutate(rc_age = -8033 * log(fm_corr),
         sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
```

```{r, eval = FALSE}
ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

```{r}
# No outliers, vs depth
combined_results %>% 
  filter(abs(fm_diff_mean) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Core depth (cm)",
       y = "Fm - mean Fm")
```

```{r, eval = FALSE}
# No outliers, vs mean_fm
combined_results %>% 
  filter(abs(fm_diff_mean) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_mean, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  geom_smooth(method = "lm", aes(fm_mean, fm_diff, color = method)) +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Mean Fm",
       y = "Fm - mean Fm")
```

@@

```{r}
combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
ggplot(aes(depth, rc_age, color = method)) +
  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
                      ymax = rc_age + sig_rc_age),
                  size = 0.5) +
  geom_smooth(alpha = .2) +
  coord_flip() +
  scale_y_reverse() +
  scale_x_reverse() +
  labs(title = "Age vs depth for graphite and HGIS",
       y = "Radiocarbon years (BP)",
       x = "Core depth (cm)")
```

@@

```{r}
plot(cal_dates_gr)
```

```{r}
plot(cal_dates_gas)
```

>>>>

## Ion currents and method blank

@@

### Ion current vs CO2 flow

@@

```{r cur-flow-os, fig.height=4, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  labs(title = "Ion current is stable for a range of CO2 flow",
       subtitle = "Current vs. CO2 flow during vial dillution",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic()
```

@@

```{r blanks, fig.height=4, dev='svg'}
cons_df %>% 
  filter(fm_consensus < 0.1) %>% 
  ggplot(aes(he12C, norm_ratio)) +
  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio  + sig_norm_ratio)) +
  geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C  + sig_he12C)) +
  geom_point(aes(color = wheel), size = 3) +
  #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
  labs(title = "Contaminant carbon is related to current",
       subtitle = "Fm and current of radiocarbon-free carbonates",
       x = "12C Current (μA)",
       y = "Fraction modern") +
  theme_classic() +
  theme(legend.position = c(0.82, 0.76),
        legend.background = element_rect(fill = "white", color = "black")) 
```

## Consensus standards

@@

cons_df %>% 
ggplot(aes(fm_consensus, fm_diff, color = Name)) + 
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = mean_diff, color = "blue") +
  geom_hline(yintercept = mean_diff + sd_diff, color = "lightblue") +
  geom_hline(yintercept = mean_diff - sd_diff, color = "lightblue") +
  geom_smooth(method = "lm",  se = FALSE) + 
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr), 
                  position = position_dodge2(width = 0.1), 
                  size = 0.2) + 
  labs(title = "Agreement of HGIS with graphite measurements",
       subtitle = "Blank corrected HGIS measurements of consensus carbonate standards",
       x = "Fm expected (points jittered for clarity)", 
       y = "Fm difference (HGIS - expected)")
```

@@

@@

combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
  mutate(rc_age = -8033 * log(fm_corr),
         sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))

# No outliers
cr_no <- combined_results %>% 
  filter(abs(fm_diff_mean) < 0.02)

mean_diff <- cr_no %>% 
  filter(method == "hgis") %>% 
  summarise(across(fm_diff, list(mean = mean, sd = sd)))

mean_errs <- cr_no %>% 
  group_by(method) %>% 
  summarise(mean_sig = mean(sig_fm_corr)) %>% 
  pivot_wider(names_from = method,
              values_from = mean_sig)
```

To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method compared to the same samples measured using the traditional graphite method. We found the measurements to be in good agreement, with no significant bias and errors averaging Fm `r mean_errs$hgis` compared with Fm `r mean_errs$graphite` for graphite measurements.

```{r}
ggplot(cr_no) +
  geom_hline(yintercept = 0) +
  #geom_hline(yintercept = mean_diff$fm_diff_mean, color = "blue") +
  #geom_hline(yintercept = mean_diff$fm_diff_mean + mean_diff$fm_diff_sd, color = "lightblue") +
  #geom_hline(yintercept = mean_diff$fm_diff_mean - mean_diff$fm_diff_sd, color = "lightblue") +
  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "HGIS measurements agree with graphite",
       subtitle = "Sediment core macrofossils measured via HGIS and graphite",
       x = "Core depth (cm)",
       y = "HGIS Fm - graphite Fm") +
  theme(legend.position = c(0.80, 0.85),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "white", color = "black")) 
```

@@

```{r}
combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
ggplot(aes(depth, rc_age, color = method, shape = method)) +
  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
                      ymax = rc_age + sig_rc_age),
                  size = 0.5) +
  geom_smooth(alpha = .2) +
  coord_flip() +
  scale_y_reverse() +
  scale_x_reverse() +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "HGIS age-depth is similar to graphite",
       subtitle = "Age vs. depth for graphite and HGIS",
       y = "Radiocarbon years (BP)",
       x = "Core depth (cm)") +
  theme(legend.position = c(0.85, 0.30),
        legend.background = element_rect(fill = "white", color = "black")) 
```

@@

```{r, fig.height=5}
bc_gr <- plot(cal_dates_gr) +
  xlim(8000, 0) +
  ylim(1000, 0) +
  labs(title = "Graphite measurements",
       xlab = "Age (yBP)",
       ylab = "Core depth (cm)")

bc_gas <- plot(cal_dates_gas) +
  xlim(8000, 0) +
  ylim(1000, 0) +
  labs(title = "HGIS measurements",
       xlab = "Age (yBP)",
       ylab = "Core depth (cm)")

bchron <- bc_gr / bc_gas

bchron +
  plot_annotation(title = "HGIS and graphite age models agree",
                  subtitle = "Calibrated ages vs core depth")
```
